Solar wind discontinuities spatial evolution

### 2024-01-25 index.qmd
better bib support

'Discontinuities' are discontinuous spatial changes in plasma parameters/characteristics and magnetic fields [@colburn1966].

They are observed across the heliosphere from inner heliosphere [@liu2022] to the heliosheath [@burlaga2011].

@@

![A rotational discontinuity (RD) detected at 0.126 AU](images/jgra57047-fig-0002-m.jpg){fig-alt="The properties of a rotational discontinuity (RD) detected at 0.126 AU on 28 January 2020. (a) The magnetic field magnitude and three components in Radial‐Tangential‐Normal (RTN) coordinate system, (b) the magnetic field in local magnetic normal (LMN) coordinate system, (c and d) the proton velocity in RTN and LMN coordinate systems, (e) the proton density, (f) the proton thermal pressure, the magnetic pressure and the sum, (g) the tangential components of the electric field in LMN coordinate system in the RD rest frame, (h) the position, orientation, and motion of the RD, (i–k) the distributions of BL, JM, and J · E across the RD, (l and m) the jump conditions across the RD. In (h), the cyan plane and arrow represent the RD and its motion vector, and the yellow curve shows the PSP orbit. In (j), both the colors and the curve correspond to the values of JM, and two vertical dashed lines indicate the area of |JM|>13max|JM| $\\vert {J}_{M}\\vert > \\frac{1}{3}\\text{max}\\left(\\vert {J}_{M}\\vert \\right)$ whose width is defined as the half thickness of the discontinuity. In (m), the solid lines present the actual upstream and downstream velocities, while the dashed lines show the theoretical downstream velocity to satisfy the Walén relation."}

@@

@söding2001 studied the radial distribution of discontinuities in the solar wind between 0.3 and 19 AU during solar activity minimum. @liu2021 studied the spatial evolution of the discontinuities from 0.13-0.9 AU, using measurements from the Parker Solar Probe. [Typical rate of occurrence is about 50 per day at 1AU]{.aside}

@@

    > The strong discontinuities produce a power-law spectrum in the 'inertial subrange' with a spectral index near the Kolmogorov -5/3 index. The discontinuity spectrum contains about half of the power of the full solar-wind magnetic ﬁeld over this 'inertial subrange'. [@borovsky2010]

## Motivations

Studying the radial distribution of occurrence rate, as well as the properties of solar wind discontinuities may help answer the following questions:

-   How does the discontinuities change with the radial distance from the Sun?
-   How is solar wind discontinuities formed? What is the physical mechanisms?
    -   Generated at or near the sun?
    -   Locally generated in the interplanetary space by turbulence?

Joint observations of JUNO & ARTEMIS & Other missions really provides a unique opportunity!!!

::: {#fig-juno_traj .column-page layout-ncol="2"}
![](https://www.nasa.gov/wp-content/uploads/2023/03/575573main_Juno20110727-3-43_full.jpg)

{{< video https://www.youtube.com/watch?v=sYp5p2oL51g title='Juno spacecraft trajectory animation' >}}

Juno Spacecraft Cruise Trajectory
:::

-   Five-year cruise to Jupiter from 2011 to 2016
-   One earth flyby in 2013
-   Nearly the same Heliographic latitude as Earth

To eliminate the effect of the solar wind structure, we use data from other missions (mainly at 1AU) to provide a way of normalization.

| Mission       | r \[AU\]  | $\delta t_B$        | $\delta t_{plasma}$                  | Data availability         |
|---------------|---------------|---------------|---------------|---------------|
| JUNO          | 1-5.5     | 1s averaged (64 Hz) | 1h **model** [\*](#solar-wind-model) | **2011 - 2016** - Today   |
| ARTEMIS       | 1         | 1s averaged (8 Hz)  | 1h averaged                          | 2009 - Today (solar wind) |
| STEREO-A      | 1         | 1s averaged (8 Hz)  | 1h averaged                          | 2006 - Today              |
| STEREO-B      | 1         | 1s averaged (8 Hz)  | 1h averaged                          | 2006 - 2016.09            |
| Wind          | 1         | 11 Hz               | 1h averaged                          | 1994 - 2004 -2020 - Today |
| Solar Orbiter | 0.28-0.91 |                     |                                      | 2020 - Today              |
| Ulysis        |           |                     |                                      |                           |

@@

- We use [@liu2022a] method to identify IDs, which has better compatibility for the IDs with minor field changes.
- Then the minimum variance analysis is applied to each ID event to obtain the boundary normal (LMN) coordinate and extract IDs' features.

@@

We are using [Michigan Solar WInd Model 2D (MSWIM2D)](http://csem.engin.umich.edu/MSWIM2D/), which models the solar wind propagation in 2D using the BATSRUS MHD solver. [@keebler2022]

@@

For model validation part, please see [JUNO Model Report](notebooks/analysis/20_model.ipynb).

# Results

For all results, see [results page](notebooks/results.qmd).

## Occurrence rates

[For code, see [noteboook](notebooks/analysis/01_occurence_rate.ipynb).]{.aside}

::: {#fig-ocr_time layout-ncol="2"}
![](figures/ocr/ocr_time_cleaned)

![](figures/ocr/ocr_time_N2_cleaned)

Occurrence rates with time
:::

::: {#fig-ocr_r layout-ncol="2"}
![](figures/ocr/ocr_r_cleaned)

![](figures/ocr/ocr_r_N2_cleaned)

Occurrence rates with radial distance
:::

## Properties

::: {#fig-thickness-distribution .column-page layout-ncol="2"}
![MN thickness](figures/thickness/thickness_mn_dist)

![MN current](figures/current/current_mn_dist)

![MN normalized thickness](figures/thickness/thickness_mn_N1_dist)

![MN normalized current](figures/current/current_mn_N1_dist)

Thickness and current density distribution
:::

>>>>

'Discontinuities' are discontinuous spatial changes in plasma parameters/characteristics and magnetic fields [@colburnDiscontinuitiesSolarWind1966].

They are observed across the heliosphere from inner heliosphere [@liuCategorizingMHDDiscontinuities2022] to the heliosheath [@burlagaCurrentSheetsHeliosheath2011].

@@

![A rotational discontinuity (RD) detected at 0.126 AU](images/jgra57047-fig-0002-m.jpg){fig-alt="The properties of a rotational discontinuity (RD) detected at 0.126 AU on 28 January 2020. (a) The magnetic field magnitude and three components in Radial‐Tangential‐Normal (RTN) coordinate system, (b) the magnetic field in local magnetic normal (LMN) coordinate system, (c and d) the proton velocity in RTN and LMN coordinate systems, (e) the proton density, (f) the proton thermal pressure, the magnetic pressure and the sum, (g) the tangential components of the electric field in LMN coordinate system in the RD rest frame, (h) the position, orientation, and motion of the RD, (i–k) the distributions of BL, JM, and J · E across the RD, (l and m) the jump conditions across the RD. In (h), the cyan plane and arrow represent the RD and its motion vector, and the yellow curve shows the PSP orbit. In (j), both the colors and the curve correspond to the values of JM, and two vertical dashed lines indicate the area of $\vert {J}_{M}\vert > \frac{1}{3}\text{max}\left(\vert {J}_{M}\vert \right)$ whose width is defined as the half thickness of the discontinuity. In (m), the solid lines present the actual upstream and downstream velocities, while the dashed lines show the theoretical downstream velocity to satisfy the Walén relation."}

@@

@sodingRadialLatitudinalDependencies2001 studied the radial distribution of discontinuities in the solar wind between 0.3 and 19 AU during solar activity minimum. @liuCharacteristicsInterplanetaryDiscontinuities2021 studied the spatial evolution of the discontinuities from 0.13-0.9 AU, using measurements from the Parker Solar Probe. [Typical rate of occurrence is about 50 per day at 1AU]{.aside}

@@

    > The strong discontinuities produce a power-law spectrum in the 'inertial subrange' with a spectral index near the Kolmogorov -5/3 index. The discontinuity spectrum contains about half of the power of the full solar-wind magnetic ﬁeld over this 'inertial subrange'. [@borovskyContributionStrongDiscontinuities2010]

@@

- We use [@liuMagneticDiscontinuitiesSolar2022] method to identify IDs, which has better compatibility for the IDs with minor field changes.
- Then the minimum variance analysis is applied to each ID event to obtain the boundary normal (LMN) coordinate and extract IDs' features.

@@

We are using [Michigan Solar WInd Model 2D (MSWIM2D)](http://csem.engin.umich.edu/MSWIM2D/), which models the solar wind propagation in 2D using the BATSRUS MHD solver. @keeblerMSWIM2DTwodimensionalOuter2022

@@

For model validation part, please see [JUNO Model Report](notebooks/20_model.ipynb).
